---
title: "DA420_project3_Grahn"
author: "Jason Grahn"
date: "1/25/2019"
output: word_document
---

#Part 1
Code is below. Output will be attached separately.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
# R preliminaries to get the user-defined function for spine chart: 
# place the spine chart code file <R_utility_program_1.R>
# in your working directory and execute it by
#     source("R_utility_program_1.R")
# Or if you have the R binary file in your working directory, use

load(file="mtpa_spine_chart.Rdata")
# spine chart accommodates up to 45 part-worths on one page
# |part-worth| <= 40 can be plotted directly on the spine chart
# |part-worths| > 40 can be accommodated through standardization

library(tibble)
library(dplyr)
```

```{r message=FALSE}
# Traditional Conjoint Analysis (R)

print.digits <- 2  # set number of digits on print and spine chart
library(support.CEs)  # package for survey construction 

# generate a balanced set of product profiles for survey
provider.survey <- Lma.design(attribute.names = 
  list(brand = c("AT&T","T-Mobile","US Cellular","Verizon"), 
  startup = c("$100","$200","$300","$400"), 
  monthly = c("$100","$200","$300","$400"),
  service = c("4G NO","4G YES"), 
  retail = c("Retail NO","Retail YES"),
  apple = c("Apple NO","Apple YES"), 
  samsung = c("Samsung NO","Samsung YES"), 
  google = c("Nexus NO","Nexus YES")), nalternatives = 1, nblocks=1, seed=9999)

#sink(here::here("project3/questions_for_survey.txt")  # send survey to external text file
#questionnaire(provider.survey)
#sink() # send output back to the screen

# user-defined function for plotting descriptive attribute names 
effect.name.map <- function(effect.name) { 
  if(effect.name=="brand") return("Mobile Service Provider")
  if(effect.name=="startup") return("Start-up Cost")
  if(effect.name=="monthly") return("Monthly Cost")
  if(effect.name=="service") return("Offers 4G Service")
  if(effect.name=="retail") return("Has Nearby Retail Store")
  if(effect.name=="apple") return("Sells Apple Products")
  if(effect.name=="samsung") return("Sells Samsung Products")
  if(effect.name=="google") return("Sells Google/Nexus Products")
  } 

# read in conjoint survey profiles with respondent ranks
conjoint.data.frame <- readr::read_csv(here::here("project3/mobile.csv"))

#building a randomizer to change the rankings in the mobile data.
set.seed(42)
ranklist <- c(1:16) #build a vector of 1 through 16
samplist <- sample(ranklist, #sample that vector and make it a list
                   size=16)
samplist <- tibble::tibble(samplist) #turn that list into it's own table

#using that table to redo the ranking list. 
conjoint.data.frame <- add_column(conjoint.data.frame, sample(ranklist,
                                                              size=16)) %>% 
  rename(ranking2 = "sample(ranklist, size = 16)") 
conjoint.data.frame
```

```{r}
# set up sum contrasts for effects coding as needed for conjoint analysis
options(contrasts=c("contr.sum","contr.poly"))

# main effects model specification
main.effects.model <- {ranking2 ~ brand + startup + monthly + service + 
  retail + apple + samsung + google}

# fit linear regression model using main effects only (no interaction terms)
main.effects.model.fit <- lm(main.effects.model, data=conjoint.data.frame)
print(summary(main.effects.model.fit)) 
```

```{r}
# save key list elements of the fitted model as needed for conjoint measures
conjoint.results <- 
  main.effects.model.fit[c("contrasts","xlevels","coefficients")]

conjoint.results$attributes <- names(conjoint.results$contrasts)

# compute and store part-worths in the conjoint.results list structure
part.worths <- conjoint.results$xlevels  # list of same structure as xlevels
end.index.for.coefficient <- 1  # intitialize skipping the intercept
part.worth.vector <- NULL # used for accumulation of part worths
for(index.for.attribute in seq(along=conjoint.results$contrasts)) {
  nlevels <- length(unlist(conjoint.results$xlevels[index.for.attribute]))
  begin.index.for.coefficient <- end.index.for.coefficient + 1
  end.index.for.coefficient <- begin.index.for.coefficient + nlevels -2
  last.part.worth <- -sum(conjoint.results$coefficients[
    begin.index.for.coefficient:end.index.for.coefficient])
  part.worths[index.for.attribute] <- 
    list(as.numeric(c(conjoint.results$coefficients[
      begin.index.for.coefficient:end.index.for.coefficient],
      last.part.worth)))
  part.worth.vector <- 
    c(part.worth.vector,unlist(part.worths[index.for.attribute]))    
  } 
conjoint.results$part.worths <- part.worths

# compute standardized part-worths
standardize <- function(x) {(x - mean(x)) / sd(x)}
conjoint.results$standardized.part.worths <- 
  lapply(conjoint.results$part.worths,standardize)
 
# compute and store part-worth ranges for each attribute 
part.worth.ranges <- conjoint.results$contrasts
for(index.for.attribute in seq(along=conjoint.results$contrasts)) 
  part.worth.ranges[index.for.attribute] <- 
  dist(range(conjoint.results$part.worths[index.for.attribute]))
conjoint.results$part.worth.ranges <- part.worth.ranges

sum.part.worth.ranges <- sum(as.numeric(conjoint.results$part.worth.ranges))

# compute and store importance values for each attribute 
attribute.importance <- conjoint.results$contrasts
for(index.for.attribute in seq(along=conjoint.results$contrasts)) 
  attribute.importance[index.for.attribute] <- 
  (dist(range(conjoint.results$part.worths[index.for.attribute]))/
  sum.part.worth.ranges) * 100
conjoint.results$attribute.importance <- attribute.importance
 
# data frame for ordering attribute names
attribute.name <- names(conjoint.results$contrasts)
attribute.importance <- as.numeric(attribute.importance)
temp.frame <- data.frame(attribute.name,attribute.importance)
conjoint.results$ordered.attributes <- 
  as.character(temp.frame[sort.list(
  temp.frame$attribute.importance,decreasing = TRUE),"attribute.name"])

# respondent internal consistency added to list structure
conjoint.results$internal.consistency <- summary(main.effects.model.fit)$r.squared 
 
# user-defined function for printing conjoint measures
if (print.digits == 2) 
  pretty.print <- function(x) {sprintf("%1.2f",round(x,digits = 2))} 
if (print.digits == 3) 
  pretty.print <- function(x) {sprintf("%1.3f",round(x,digits = 3))} 
 
# report conjoint measures to console 
# use pretty.print to provide nicely formated output
for(k in seq(along=conjoint.results$ordered.attributes)) {
  cat("\n","\n")
  cat(conjoint.results$ordered.attributes[k],"Levels: ",
  unlist(conjoint.results$xlevels[conjoint.results$ordered.attributes[k]]))
  
  cat("\n"," Part-Worths:  ")
  cat(pretty.print(unlist(conjoint.results$part.worths
    [conjoint.results$ordered.attributes[k]])))
    
  cat("\n"," Standardized Part-Worths:  ")
  cat(pretty.print(unlist(conjoint.results$standardized.part.worths
    [conjoint.results$ordered.attributes[k]])))  
    
  cat("\n"," Attribute Importance:  ")
  cat(pretty.print(unlist(conjoint.results$attribute.importance
    [conjoint.results$ordered.attributes[k]])))
  }

# plotting of spine chart begins here
# all graphical output is routed to external pdf file
pdf(file = here::here("project3/fig_preference_mobile_services_results.pdf"), width=8.5, height=11)
spine.chart(conjoint.results)
dev.off()  # close the graphics output device

# Suggestions for the student:
# Enter your own rankings for the product profiles and generate
# conjoint measures of attribute importance and level part-worths.
# Note that the model fit to the data is a linear main-effects model.
# See if you can build a model with interaction effects for service
# provider attributes.
```


#Part 2

Consider Exhibit 4.1 in page 50-51. Suppose your client is someone other than the local farmer, a meat producer/butcher, dairy, or brewer perhaps. Determine association rules relevant to your client’s products guided by the market basket model. What recommendations would you make about future marketplace actions?

## Analysis
We selected running the market basket analysis for a brewer, and opted to find rules related to "beer". 

The results included 6 rules, the strongest of which was for non-alcoholic beer. That is to say, when a consumner purchases non-alcoholic beer, they will also purchase beer. The second highest rule is simply beer by itself. The remaining rules in order are bread and baked goods, vegetables, fruit, and dairy produce.

Given these rules, it makes sense that the brewer investigate producing a non-alcoholic beer of their own if they haven't already. Given the high association of non-alcoholic beer with beer, consumer confidence in a non-alcoholic beer could lead to purchases of the same companies' alcoholic beer as well, as "birds of a feather." The remaining rules indicate purchases of items used for cookouts (bread/baked goods, veggies, fruit, and dairy produce). Dairy produce in this case could be cheeses, while bread and baked goods could related to hotdog and hamburger buns. 

Additionally, it makes market-sense for the brewer to align their product with a "friendly" supplier of breads and dairy. Perhaps they might offer coupons or discounts to select grocery stores; when customers purchase a these associated products, they receive a discount on beer. As an alternative, purchasing beer of their brand could enable discounts on select breads and baked products.

Given the relationship of beer to these items, we must also inquire toward the seasonality of the data-set. This merits additional study to determine if decisions made now might only be good for a small window of time during the year. 

```{r}
# Association Rules for Market Basket Analysis (R)

library(arules)  # association rules
library(arulesViz)  # data visualization of association rules
library(RColorBrewer)  # color palettes for plots

data(Groceries)  # grocery transactions object from arules package

# examine frequency for each item with support greater than 0.025
itemFrequencyPlot(Groceries, support = 0.025, cex.names=0.8, xlim = c(0,0.3),
                  type = "relative", horiz = TRUE, col = "light blue", las = 1,
                  xlab = paste("Proportion of Market Baskets Containing Item",
                               "\n(Item Relative Frequency or Support)"))
```

```{r}
# explore possibilities for combining similar items
print(head(itemInfo(Groceries))) 
print(levels(itemInfo(Groceries)[["level1"]]))  # 10 levels... too few 
print(levels(itemInfo(Groceries)[["level2"]]))  # 55 distinct levels
```

```{r}
# aggregate items using the 55 level2 levels for food categories
# to create a more meaningful set of items
groceries <- aggregate(Groceries, itemInfo(Groceries)[["level2"]])  

#print(dim(groceries)[1])  # 9835 market baskets for shopping trips
#print(dim(groceries)[2])  # 55 final store items (categories)  
itemFrequencyPlot(groceries, support = 0.025, cex.names=1.0, xlim = c(0,0.5),
                  type = "relative", horiz = TRUE, col = "dark blue", las = 1,
                  xlab = paste("Proportion of Market Baskets Containing Item",
                               "\n(Item Relative Frequency or Support)"))
```


```{r message=FALSE}
# obtain large set of association rules for items by category and all shoppers
# this is done by setting very low criteria for support and confidence
first.rules <- apriori(groceries, parameter = list(support = 0.001, confidence = 0.05))

# select association rules using thresholds for support and confidence 
second.rules <- apriori(groceries, parameter = list(support = 0.025, confidence = 0.05))

# data visualization of association rules in scatter plot
plot(second.rules, 
     control=list(jitter=2, 
                  col = rev(brewer.pal(9, "Greens")[4:9])),
     shading = "lift")
```

```{r}
# grouped matrix of rules is too clustered to show anything
plot(second.rules, method="grouped",   
  control=list(col = rev(brewer.pal(9, "Greens")[4:9])))
```

```{r}
# select rules with vegetables in consequent (right-hand-side) item subsets
beer.rules <- subset(second.rules, subset = rhs %pin% "beer")
inspect(beer.rules)  # 6 rules
```


```{r}
# sort by lift and identify the top 10 rules
top.beer.rules <- head(sort(beer.rules, decreasing = TRUE, by = "lift"), 10)
inspect(top.beer.rules) 
```

Or visually:
```{r}
plot(top.beer.rules, method="grouped",   
  control=list(col = rev(brewer.pal(9, "Greens")[4:9])))
```

```{r}
plot(top.beer.rules, method="graph", 
  control=list(type="items"), 
  shading = "lift")
```



